---
title: "draft_report"
author: "Hannah, Jeff, Josh Lin, Pauline"
date: "11/16/2020"
output: pdf_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../reports/") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(ggplot2)
library(sandwich)
library(lmtest)
library(stargazer)
library(patchwork)
```

```{r, warning = FALSE}
#combining the two excel sheets 
d <- read_excel("../data/interim/covid-19_v2.xlsx", sheet = "Covid-19", skip=1)
reopen <- read_excel("../data/interim/reopen_CUSP.xlsx", sheet = "Reopening")
total <- merge(d,reopen, by="state")
```

```{r}
#make values numeric 
total$reopen_state_bus = as.numeric(total$reopen_state_bus)
total$reopen_business_statewide = as.numeric(total$reopen_business_statewide)
```


```{r}
total$shelter_in_place = excel_numeric_to_date(total$shelter_in_place)
total$end_shelter_in_place = excel_numeric_to_date(total$end_shelter_in_place)
total$close_nonessential_business = excel_numeric_to_date(total$close_nonessential_business)
total$reopen_business_statewide = excel_numeric_to_date(total$reopen_business_statewide)
total$mandate_facemask = excel_numeric_to_date(total$mandate_facemask)
total$mandate_facemask_employee = excel_numeric_to_date(total$mandate_facemask_employee)
total$end_statewide_mandate_facemask = excel_numeric_to_date(total$end_statewide_mandate_facemask)

total$reopen_restaurants = excel_numeric_to_date(total$reopen_restaurants)
total$reopen_childcare = excel_numeric_to_date(total$reopen_childcare)
total$reopen_business_statewide = excel_numeric_to_date(as.numeric(total$reopen_business_statewide))
total$reopen_gyms = excel_numeric_to_date(total$reopen_gyms)
total$reopen_hair_salons = excel_numeric_to_date(total$reopen_hair_salons)
total$reopen_movie_theaters = excel_numeric_to_date(total$reopen_movie_theaters)
total$`reopen_non-essential_retail` = excel_numeric_to_date(total$`reopen_non-essential_retail`)
total$reopen_religious = excel_numeric_to_date(total$reopen_religious)
total$reopen_state_bus = excel_numeric_to_date(as.numeric(total$reopen_state_bus))
total$reopen_bars = excel_numeric_to_date(total$reopen_bars)
```

```{r}
write_csv(total, file = "../data/interim/covid19+cusp.csv")
```

```{r}
plot(total$`length_ shelter_in_place`, total$cases_per_100k)
```

```{r}
#11/16 baseline model re: current research q

party_model <- lm(cases_per_100k ~ `length_ shelter_in_place` + political_party_governor, data=x)
summary(party_model)

```
```{r}
#filtering to find number of cases in dem and repub states 
dems <- filter(total, political_party_governor == 'Democratic')
repub <- filter(total, political_party_governor == 'Republican')
dem_cases <- sum(dems$cases_per_100k)
repub_cases <- sum(repub$cases_per_100k)
party_cases<- data.frame(dem_cases, repub_cases)
```


```{r}
#making a barplot of cases for total democratic and republican states 
c1 <- rgb(173,216,230,max = 255, alpha = 200, names = "lt.blue")
c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")
barplot(c(dem_cases, repub_cases), main="State Political Party and Cases",
        ylab = 'Cases per 100K',
        names.arg=c('Democratic', 'Republican'),
        col = c(c1,c2), ylim=c(0, 80000))
```



```{r}
#Improvement v1
poverty_model <- lm(cases_per_100k ~ `length_ shelter_in_place` + political_party_governor + percent_povertyline_2018 + as.numeric(white_percent_cases), data=x)
summary(poverty_model)
```




```{r}
#Improvement v2

poverty_model_v2 <- lm(cases_per_100k ~ `length_ shelter_in_place` + political_party_governor + percent_povertyline_2018 + as.numeric(white_percent_cases) + as.numeric(black_percent_cases) + as.numeric(hispanic_percent_cases) + as.numeric(other_percent_cases), data=x)
summary(poverty_model_v2)
#Due to NA's in dataset, the model only includes 43 observations for each variable. Attempts were made to find data to fill in the gaps, but no public data was found for each instance of an NA

```


```{r, warning = FALSE}

#white cases: remove NA and make values numeric 
total$white_percent_cases[total$white_percent_cases == 'NR'] <- NA
na_white_percent_cases <- na.omit(as.numeric(total$white_percent_cases))

#black cases: remove NA, make values numeric, replace value of '<0.1' with '0.007' from external data *state = Montana
total$black_percent_cases[total$black_percent_cases == '<0.1'] <- '0.007'
total$black_percent_cases[total$black_percent_cases == 'NR'] <- NA
na_black_percent_cases <- (na.omit(as.numeric(total$black_percent_cases)))

#hispanic cases: remove NA and make values numeric
total$hispanic_percent_cases[total$hispanic_percent_cases == 'NR'] <- NA
na_hispanic_percent_cases <- na.omit(as.numeric(total$hispanic_percent_cases))

#other cases: remove Na, make values numeric, replace value of '<0.1' with '0.02' from external data *state = Texas, note = other from external data includes Asian ethnicity only  
total$other_percent_cases[total$other_percent_cases == '<0.1'] <- '0.02'
total$other_percent_cases[total$other_percent_cases == 'NR'] <- NA
na_other_percent_cases <- na.omit(as.numeric(total$other_percent_cases)) 


#sum number of cases for each ethnicity 
white <- sum(total$total_cases*as.numeric(na_white_percent_cases))
black <- sum(total$total_cases*as.numeric(na_black_percent_cases))
hispanic <- sum(total$total_cases*as.numeric(na_hispanic_percent_cases))
other <- sum(total$total_cases*as.numeric(na_other_percent_cases))

```



```{r}
#making a barplot of cases for cases by ethnicity 
c1 <- rgb(173,216,230,max = 255, alpha = 200, names = "lt.blue")
c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")

barplot(c(white, black, hispanic, other), main="Cases by Ethnicity",
        ylab = 'Total Cases',
        names.arg=c('White', 'Black', 'Hispanic', 'Other'),
        col = c(c1,c2,"palegreen3", "yellow"), ylim=c(0, 5000000))
```


```{r, warning=FALSE}
se.model1 = coeftest(party_model, vcov = vcovHC)[ , "Std. Error"]
se.model2 = coeftest(poverty_model, vcov = vcovHC)[ , "Std. Error"]
se.model3 = coeftest(poverty_model_v2, vcov = vcovHC)[ , "Std. Error"]

stargazer(party_model, poverty_model, poverty_model_v2,
          type = "text", omit.stat = "f", se = list(se.model3),
          star.cutoffs = c(0.05, 0.01, 0.001),
          model.numbers = FALSE,
          column.labels = c('Baseline', 'Improvement (v1)', 'Improvement (v2)'),
          dep.var.labels = c('Cases per 100000 Individuals'),
          covariate.labels = c('Sheltering in Place Period', 'Political Party (Republican)', 'Percent under Poverty Line',
                              'Percent Covid Cases (White)', "Percent Covid Cases (Black)",
                              'Percent Covid Cases (Hispanic)', 'Percent Covid Cases (Other)'),
          title = "Table 1: Relationship between cases per 100000 and state political party")
```

###CLM Assumptions 

##IID Sampling (Assumption 1)
The data is IID because the reference population sampled was all the people testing for COVID-19 in the United States. It is difficult to determine if there is clustering in the data because we do not have access to data for each individual who tested for COVID-19. However because our research question is concerning state level policies and characteristics we believe that our data is suitable to answer our research question. 

# Linear Conditional Expectation (Assumption 2)

```{r}
x <- total[-c(12, 19, 27, 32, 33, 35, 44, 49), ]
```



```{r}

x_resid <- x %>%  
  mutate(party_model_resid = resid(party_model),
         party_model_prediction = predict(party_model),
         poverty_model_resid = resid(poverty_model),
         poverty_model_prediction = predict(poverty_model),
         poverty_model_v2_resid = resid(poverty_model_v2),
         poverty_model_v2_prediction = predict(poverty_model_v2))

plot_1 <- x_resid %>%  
  ggplot(aes(x = party_model_prediction, y = party_model_resid)) + 
  geom_point() + stat_smooth()

plot_2 <- x_resid %>%  
  ggplot(aes(x = poverty_model_prediction, y = poverty_model_resid)) + 
  geom_point() + stat_smooth()

plot_3 <- x_resid %>%  
  ggplot(aes(x = poverty_model_v2_prediction, y = poverty_model_v2_resid)) + 
  geom_point() + stat_smooth()

```

```{r} 
plot_1
```


```{r}
plot_2
```



```{r}
plot_3
```

## Consequences of a non-linear conditional expectation

The model you have fitted that assumes the data is linear and the model you have fitted is still the best **linear** predictor. However, the estimated coefficient does not match the relationship in the data. This means that: 
  1. There are other families of nonlinear models that are more efficient
  2. Linear relationships don't *fully* model the complexity of the data
  3. Prediction (and inference) can be improved with more complexity
  
  
```{r}
summary(party_model)
summary(poverty_model)
summary(poverty_model_v2)
```


```{r}
poverty_model_v2
```
  

```{r}
summary(lm(cases_per_100k ~ `length_ shelter_in_place`, data = x))
  
summary(lm(cases_per_100k ~ political_party_governor, data = x))

summary(lm(cases_per_100k ~ percent_povertyline_2018, data = x))

summary(lm(cases_per_100k ~ as.numeric(white_percent_cases), data = x))

summary(lm(cases_per_100k ~ as.numeric(black_percent_cases), data = x))

summary(lm(cases_per_100k ~ as.numeric(hispanic_percent_cases), data = x))

summary(lm(cases_per_100k ~ as.numeric(other_percent_cases), data = x))

  
```


We do not have colinearity!! Woo 


# Homoskedastic Conditional Variance (Assumption 4)

Ocular test
```{r}
plot_4 <- x_resid %>%  
  ggplot(aes(x = party_model_prediction, y = party_model_resid)) + 
  geom_point() + geom_hline(yintercept = mean(x_resid$party_model_resid))

plot_5 <- x_resid %>%  
  ggplot(aes(x = poverty_model_prediction, y = poverty_model_resid)) + 
  geom_point() + geom_hline(yintercept = mean(x_resid$poverty_model_resid))

plot_6 <- x_resid %>%  
  ggplot(aes(x = poverty_model_v2_prediction, y = poverty_model_v2_resid)) + 
  geom_point() + geom_hline(yintercept = mean(x_resid$poverty_model_v2_resid))

```


```{r}
plot_4
```

```{r}
plot_5
```

```{r}
plot_6
```


We have homoskedasticity !



## Normally Distributed Errors (Assumption 5)

```{r}
hist(x_resid$party_model_resid)
```


```{r}
hist(x_resid$poverty_model_resid)
```

```{r}
hist(x_resid$poverty_model_v2_resid)
```

We have normally distributed standard errors ! 
