---
title: "draft_report"
author: "Hannah, Jeff, Josh Lin, Pauline"
date: "11/16/2020"
output: pdf_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../reports/") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(ggplot2)
library(sandwich)
library(lmtest)
library(stargazer)
library(patchwork)
```

```{r, warning = FALSE}
#combining the two excel sheets 
d <- read_excel("../data/interim/covid-19_v2.xlsx", sheet = "Covid-19", skip=1)
reopen <- read_excel("../data/interim/reopen_CUSP.xlsx", sheet = "Reopening")
total <- merge(d,reopen, by="state")
```

```{r}
#make values numeric due to formatting issues in excel
total$reopen_state_bus = as.numeric(total$reopen_state_bus)
total$reopen_business_statewide = as.numeric(total$reopen_business_statewide)
```


```{r}
#reformat numeric values as dates using janitor::excel_numeric_to_date
total$shelter_in_place = excel_numeric_to_date(total$shelter_in_place)
total$end_shelter_in_place = excel_numeric_to_date(total$end_shelter_in_place)
total$close_nonessential_business = excel_numeric_to_date(total$close_nonessential_business)
total$reopen_business_statewide = excel_numeric_to_date(total$reopen_business_statewide)
total$mandate_facemask = excel_numeric_to_date(total$mandate_facemask)
total$mandate_facemask_employee = excel_numeric_to_date(total$mandate_facemask_employee)
total$end_statewide_mandate_facemask = excel_numeric_to_date(total$end_statewide_mandate_facemask)

total$reopen_restaurants = excel_numeric_to_date(total$reopen_restaurants)
total$reopen_childcare = excel_numeric_to_date(total$reopen_childcare)
total$reopen_business_statewide = excel_numeric_to_date(as.numeric(total$reopen_business_statewide))
total$reopen_gyms = excel_numeric_to_date(total$reopen_gyms)
total$reopen_hair_salons = excel_numeric_to_date(total$reopen_hair_salons)
total$reopen_movie_theaters = excel_numeric_to_date(total$reopen_movie_theaters)
total$`reopen_non-essential_retail` = excel_numeric_to_date(total$`reopen_non-essential_retail`)
total$reopen_religious = excel_numeric_to_date(total$reopen_religious)
total$reopen_state_bus = excel_numeric_to_date(as.numeric(total$reopen_state_bus))
total$reopen_bars = excel_numeric_to_date(total$reopen_bars)
```

```{r}
#Save data as csv into github
write_csv(total, file = "../data/interim/covid19+cusp.csv")
```

```{r}
plot(total$`length_ shelter_in_place`, total$cases_per_100k)
```

```{r}
#remove states with NA values within the columns for percent cases by ethnicity (these variables are used in our third model down below)
x <- total[-c(12, 19, 32, 33, 35, 49), ]
```

```{r}
#11/16 baseline model re: current research q

party_model <- lm(cases_per_100k ~ `length_ shelter_in_place` + political_party_governor, data=x)
summary(party_model)

```
```{r}
#filtering to find number of cases in dem and repub states 
dems <- filter(total, political_party_governor == 'Democratic')
repub <- filter(total, political_party_governor == 'Republican')
dem_cases <- sum(dems$cases_per_100k)
repub_cases <- sum(repub$cases_per_100k)
party_cases<- data.frame(dem_cases, repub_cases)

```


```{r}
#making a barplot of cases for total democratic and republican states 
c1 <- rgb(173,216,230,max = 255, alpha = 200, names = "lt.blue")
c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")
barplot(c(dem_cases, repub_cases), main="State Political Party and Cases",
        ylab = 'Cases per 100K',
        names.arg=c('Democratic', 'Republican'),
        col = c(c1,c2), ylim=c(0, 80000))
```



```{r}
#Improvement v1
poverty_model <- lm(cases_per_100k ~ `length_ shelter_in_place` + political_party_governor + percent_povertyline_2018 + as.numeric(white_percent_cases), data=x)
summary(poverty_model)
```




```{r}
#Improvement v2

poverty_model_v2 <- lm(cases_per_100k ~ `length_ shelter_in_place` + political_party_governor + percent_povertyline_2018 + as.numeric(white_percent_cases) + as.numeric(black_percent_cases) + as.numeric(hispanic_percent_cases) + as.numeric(other_percent_cases), data=x)
summary(poverty_model_v2)
#Due to NA's in dataset, the model only includes 45 observations for each variable. Attempts were made to find data to fill in the gaps, but no public data was found for each instance, except for Montana and Texas

```


```{r, warning = FALSE}

#white cases: remove NA and make values numeric 
total$white_percent_cases[total$white_percent_cases == 'NR'] <- NA
na_white_percent_cases <- na.omit(as.numeric(total$white_percent_cases))

#black cases: remove NA, make values numeric, replace value of '<0.1' with '0.007' from external data *state = Montana
total$black_percent_cases[total$black_percent_cases == '<0.1'] <- '0.007'
total$black_percent_cases[total$black_percent_cases == 'NR'] <- NA
na_black_percent_cases <- (na.omit(as.numeric(total$black_percent_cases)))

#hispanic cases: remove NA and make values numeric
total$hispanic_percent_cases[total$hispanic_percent_cases == 'NR'] <- NA
na_hispanic_percent_cases <- na.omit(as.numeric(total$hispanic_percent_cases))

#other cases: remove Na, make values numeric, replace value of '<0.1' with '0.02' from external data *state = Texas, note = other from external data includes Asian ethnicity only  
total$other_percent_cases[total$other_percent_cases == '<0.1'] <- '0.02'
total$other_percent_cases[total$other_percent_cases == 'NR'] <- NA
na_other_percent_cases <- na.omit(as.numeric(total$other_percent_cases)) 


#sum number of cases for each ethnicity 
white <- sum(total$total_cases*as.numeric(na_white_percent_cases))
black <- sum(total$total_cases*as.numeric(na_black_percent_cases))
hispanic <- sum(total$total_cases*as.numeric(na_hispanic_percent_cases))
other <- sum(total$total_cases*as.numeric(na_other_percent_cases))
```



```{r}
#making a barplot of cases for cases by ethnicity 
c1 <- rgb(173,216,230,max = 255, alpha = 200, names = "lt.blue")
c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")

barplot(c(white, black, hispanic, other), main="Cases by Ethnicity",
        ylab = 'Total Cases',
        names.arg=c('White', 'Black', 'Hispanic', 'Other'),
        col = c(c1,c2,"palegreen3", "yellow"), ylim=c(0, 5000000))
```


```{r, warning=FALSE}
se.model1 = coeftest(party_model, vcov = vcovHC)[ , "Std. Error"]
se.model2 = coeftest(poverty_model, vcov = vcovHC)[ , "Std. Error"]
se.model3 = coeftest(poverty_model_v2, vcov = vcovHC)[ , "Std. Error"]

stargazer(party_model, poverty_model, poverty_model_v2,
          type = "text", omit.stat = "f", se = list(se.model3),
          star.cutoffs = c(0.05, 0.01, 0.001),
          model.numbers = FALSE,
          column.labels = c('Baseline', 'Improvement (v1)', 'Improvement (v2)'),
          dep.var.labels = c('Cases per 100000 Individuals'),
          covariate.labels = c('Sheltering in Place Period', 'Political Party (Republican)', 'Percent under Poverty Line',
                              'Percent Covid Cases (White)', "Percent Covid Cases (Black)",
                              'Percent Covid Cases (Hispanic)', 'Percent Covid Cases (Other)'),
          title = "Table 1: Relationship between cases per 100000 and state political party")
```

###CLM Assumptions 

##IID Sampling (Assumption 1)
The data is IID because the reference population sampled was all the people testing for COVID-19 in the United States. It is difficult to determine if there is clustering in the data because we do not have access to data for each individual who tested for COVID-19. However because our research question is concerning state level policies and characteristics we believe that our data is suitable to answer our research question. 

# Linear Conditional Expectation (Assumption 2)


```{r}

x_resid <- x %>%  
  mutate(party_model_resid = resid(party_model),
         party_model_prediction = predict(party_model),
         poverty_model_resid = resid(poverty_model),
         poverty_model_prediction = predict(poverty_model),
         poverty_model_v2_resid = resid(poverty_model_v2),
         poverty_model_v2_prediction = predict(poverty_model_v2))

model1_plot_1 <- x_resid %>%  
  ggplot(aes(x = `length_ shelter_in_place`, y = party_model_resid)) + 
  geom_point() + stat_smooth()

model1_plot_2 <- x_resid %>%  
  ggplot(aes(x = `political_party_governor`, y = party_model_resid)) + 
  geom_point() + stat_smooth()

model1_plot_3 <- x_resid %>%  
  ggplot(aes(x = party_model_prediction, y = party_model_resid)) + 
  geom_point() + stat_smooth()

```

```{r}
model1_plot_1 | model1_plot_2 | model1_plot_3
```

```{r}

model2_plot_1 <- x_resid %>%  
  ggplot(aes(x = `length_ shelter_in_place`, y = poverty_model_resid)) + 
  geom_point() + stat_smooth()

model2_plot_2 <- x_resid %>%  
  ggplot(aes(x = `political_party_governor`, y = poverty_model_resid)) + 
  geom_point() + stat_smooth()

model2_plot_3 <- x_resid %>%  
  ggplot(aes(x = percent_povertyline_2018, y = poverty_model_resid)) + 
  geom_point() + stat_smooth()

model2_plot_4 <- x_resid %>%  
  ggplot(aes(x = as.numeric(white_percent_cases), y = poverty_model_resid)) + 
  geom_point() + stat_smooth()

model2_plot_5 <- x_resid %>%  
  ggplot(aes(x = poverty_model_prediction, y = poverty_model_resid)) + 
  geom_point() + stat_smooth()

```

```{r}
model2_plot_1 | model2_plot_2
model2_plot_3 | model2_plot_4
model2_plot_5
```


```{r}
model3_plot_1 <- x_resid %>%  
  ggplot(aes(x = `length_ shelter_in_place`, y = poverty_model_v2_resid)) + 
  geom_point() + stat_smooth()

model3_plot_2 <- x_resid %>%  
  ggplot(aes(x = `political_party_governor`, y = poverty_model_v2_resid)) + 
  geom_point() + stat_smooth()

model3_plot_3 <- x_resid %>%  
  ggplot(aes(x = percent_povertyline_2018, y = poverty_model_v2_resid)) + 
  geom_point() + stat_smooth()

model3_plot_4 <- x_resid %>%  
  ggplot(aes(x = as.numeric(white_percent_cases), y = poverty_model_v2_resid)) + 
  geom_point() + stat_smooth()

model3_plot_5 <- x_resid %>%  
  ggplot(aes(x = as.numeric(black_percent_cases), y = poverty_model_v2_resid)) + 
  geom_point() + stat_smooth()
  
model3_plot_6 <- x_resid %>%  
  ggplot(aes(x = as.numeric(hispanic_percent_cases), y = poverty_model_v2_resid)) + 
  geom_point() + stat_smooth()

model3_plot_7 <- x_resid %>%  
  ggplot(aes(x = as.numeric(other_percent_cases), y = poverty_model_v2_resid)) + 
  geom_point() + stat_smooth()

model3_plot_8 <- x_resid %>%  
  ggplot(aes(x = poverty_model_v2_prediction, y = poverty_model_v2_resid)) + 
  geom_point() + stat_smooth()

```

```{r}
model3_plot_1 | model3_plot_2
model3_plot_3 | model3_plot_4
model3_plot_5 | model3_plot_6
model3_plot_7 | model3_plot_8
```

## Consequences of a non-linear conditional expectation

The model you have fitted that assumes the data is linear and the model you have fitted is still the best **linear** predictor. However, the estimated coefficient does not match the relationship in the data. This means that: 
  1. There are other families of nonlinear models that are more efficient
  2. Linear relationships don't *fully* model the complexity of the data
  3. Prediction (and inference) can be improved with more complexity
  
  
```{r}
summary(party_model)
summary(poverty_model)
summary(poverty_model_v2)
```

  

```{r}
summary(lm(cases_per_100k ~ `length_ shelter_in_place`, data = x))
  
summary(lm(cases_per_100k ~ political_party_governor, data = x))

summary(lm(cases_per_100k ~ percent_povertyline_2018, data = x))

summary(lm(cases_per_100k ~ as.numeric(white_percent_cases), data = x))

summary(lm(cases_per_100k ~ as.numeric(black_percent_cases), data = x))

summary(lm(cases_per_100k ~ as.numeric(hispanic_percent_cases), data = x))

summary(lm(cases_per_100k ~ as.numeric(other_percent_cases), data = x))

  
```


We do not have colinearity!! Woo 


# Homoskedastic Conditional Variance (Assumption 4)

Ocular test
```{r}
plot_4 <- x_resid %>%  
  ggplot(aes(x = party_model_prediction, y = party_model_resid)) + 
  geom_point() + geom_hline(yintercept = mean(x_resid$party_model_resid))

plot_5 <- x_resid %>%  
  ggplot(aes(x = poverty_model_prediction, y = poverty_model_resid)) + 
  geom_point() + geom_hline(yintercept = mean(x_resid$poverty_model_resid))

plot_6 <- x_resid %>%  
  ggplot(aes(x = poverty_model_v2_prediction, y = poverty_model_v2_resid)) + 
  geom_point() + geom_hline(yintercept = mean(x_resid$poverty_model_v2_resid))

```


```{r}
plot_4
```

```{r}
plot_5
```

```{r}
plot_6
```


We have homoskedasticity !



## Normally Distributed Errors (Assumption 5)

```{r}
hist(x_resid$party_model_resid, main = "Histogram of Baseline Model Residuals")
```

```{r}
qqnorm(x_resid$party_model_resid, main = "Q-Q Plot for Baseline Model Residuals")
qqline(x_resid$party_model_resid, col = "steelblue", lwd = 2)
```


```{r}
hist(x_resid$poverty_model_resid, main = "Histogram of Improvement(v1) Model Residuals")
```

```{r}
qqnorm(x_resid$poverty_model_resid, main = "Q-Q Plot for Improvement (v1) Model Residuals")
qqline(x_resid$poverty_model_resid, col = "steelblue", lwd = 2)
```

```{r}
#Histogram for Improvement (v2) Model 
hist(x_resid$poverty_model_v2_resid, main = "Histogram of Improvement (v2) Model Residuals")
```

```{r}
#QQ Plot for Improvement (v2) Model 
qqnorm(x_resid$poverty_model_v2_resid, main = "Q-Q Plot for Improvement (v2) Model Residuals")
qqline(x_resid$poverty_model_v2_resid, col = "steelblue", lwd = 2)
```

We have normally distributed standard errors ! 
